Incorporating Gyroscope Information in Gesture Recognition

### 2011-12-17 paper/paper.tex
editing

On a pragmatic note, we develop and test on the iPhone platform, although our solutions have general applicability, and an available Matlab implementation. The iPhone is chosen amongst our team as two of our members have previous iPhone development experience.  
\pagebreak

@@

KHMM uses k-means clustering to discretize the space of real vectors followed by training one Hidden Markov Model for each gesture type.  The intuition behind this can be thought of as follows: the clustering across all gesture types serves mainly to project each data vector which is in $ \mathbb{R}^n $ (where there are $n$ features) to $\mathbb{Z^{+}}$.  Concretely this discretizes the data uniformly regardless of what gesture type it belongs to.  Following the discretization KHMM then trains a Hidden Markov Model of the discretized training sets for each gesture type.
\\
\\
Formally we describe the algorithm as follows:

	\subsubsection{Clustering}
		For all training example across all gesture types, kmeans defines a clustering function $f: \mathbb{R}^n \rightarrow \mathbb{Z^{+}}$. 
	\subsubsection{Hidden Markov Model}
		For each set of training examples for each gesture type train a Hidden Markov Model.  Namely in this stage, for each gesture type $g$, we generate a function $h_g: (\mathbb{Z^{+}}^k) \rightarrow \mathbb{R}_{[0,1]}$  by the Hidden Markov Model process:
$HMM: (\mathbb{Z^{+}}^k)^t \rightarrow h_g$, where each training example consists of $k$ instantaneous data points and there are $t$ training examples for gesture type g.  
	\subsubsection{Classification Procedure}
		To classify an unknown gesture example, which exists in $(\mathbb{R}^n)^k$, we first transform it to its discretized version, namely $f(\mathbb{R}^n)^k$ and then for each gesture type $g$, apply $h_g((f(\mathbb{R}^n))^k)$ and classify the example according to the $g’$ were $h_g’$ returns maximum value.

@@

The motivation for Cluster-Matching-HMM derives from the realization that in the clustering step KHMM applies the same centroids to each training example regardless of what gesture type it belongs too.  We hypothesize that different gesture types have vastly different vector features score (ie lie in different subspaces of $\mathbb{R}^n$) and thus we might gain information by clustering each gesture type independently and then training an HMM for each clustered training type.  In the classification step we then classify over each clustering and each corresponding HMM.  Moreover we weight the resulting HMM probabilities according to how well the training example ``fits" in the the assigned cluster.  
\\
\\
Formally we describe the algorithm as follows:

	\subsubsection{Clustering}
		For each gesture type $g$ across all training examples, kmeans defines a clustering function $f_g: \mathbb{R}^n \rightarrow \mathbb{Z^{+}}$.
	\subsubsection{Hidden Markov Model}
		For each set of training examples for each gesture type train a Hidden Markov Model.  Namely in this set we generate a function $h_g: (\mathbb{Z^{+}}^k) \rightarrow \mathbb{R}_{[0,1]}$ by the Hidden Markov Model process: 
		$HMM: (\mathbb{Z^{+}}^k)^t \rightarrow h_g$, where each training example consists of $k$ instantaneous data points and there are $t$ training examples for gesture type $g$.

@@

		\subsubsection{Clustering}
		For each gesture type $g$ across all training examples, kmeans defines a clustering function $f_g: \mathbb{R}^n \rightarrow \mathbb{Z^{+}}$.

@@

FINISH THIS

>>>>

On a pragmatic note, we develop and test on the iPhone platform, although our solutions have general applicability, and an available Matlab implementation. The iPhone is chosen amongst our team as two of our members have previous iPhone development experience.  
% \pagebreak

@@

KHMM uses k-means clustering to discretize the space of real vectors followed by training one Hidden Markov Model for each gesture type.  The intuition behind this can be thought of as follows: the clustering across all gesture types serves mainly to project each data vector which is in $ \mathbb{R}^n $ (where there are $n$ features) to $\mathbb{Z^{+}}$.  Concretely, this discretizes the data uniformly regardless of what gesture type it belongs to.  Following the discretization KHMM then trains a Hidden Markov Model of the discretized training sets for each gesture type.
\\
\\
Formally we describe the algorithm as follows:

	\subsubsection{Clustering}
		For all training examples across all gesture types, running k-means on all of the training data simultaneously defines a single clustering function $f: \mathbb{R}^n \rightarrow \mathbb{Z^{+}}$. 
	\subsubsection{Hidden Markov Model}
		For each set of training examples for each gesture type, train a Hidden Markov Model.  Namely in this stage, for each gesture type $g$, we generate a function $h_g: (\mathbb{Z^{+}}^k) \rightarrow \mathbb{R}_{[0,1]}$  by the Hidden Markov Model process:
$HMM: (\mathbb{Z^{+}}^k)^t \rightarrow h_g$, where each training example consists of $k$ instantaneous data points and there are $t$ training examples for gesture type g.  
	\subsubsection{Classification Procedure}
		To classify an unknown gesture example, which exists in $(\mathbb{R}^n)^k$, we first transform it to its discretized version, namely $f(\mathbb{R}^n)^k$ and then for each gesture type $g$, apply $h_g((f(\mathbb{R}^n))^k)$ and classify the example according to the $g’$ were $h_g’$ returns maximum probability.

@@

The motivation for Cluster-Matching-HMM derives from the realization that in the clustering step KHMM applies the same centroids to each training example regardless of what gesture type it belongs to.  We hypothesize that different gesture types have vastly different vector features score (i.e. lie in different subspaces of $\mathbb{R}^n$) and thus we might gain information by clustering each gesture type independently and then training an HMM for each clustered training type.  In the classification step we then classify over each clustering and each corresponding HMM.  Moreover we weight the resulting HMM probabilities according to how well the training example ``fits" in the the assigned cluster.  
\\
\\
Formally we describe the algorithm as follows:

	\subsubsection{Clustering}
		For each gesture type $g$ across all training examples, running k-means on all of the training data from for one gesture only defines a clustering function $f_g: \mathbb{R}^n \rightarrow \mathbb{Z^{+}}$ for each gesture type.
	\subsubsection{Hidden Markov Model}
		For each set of training examples for each gesture type, train a Hidden Markov Model.  Namely in this set we generate a function $h_g: (\mathbb{Z^{+}}^k) \rightarrow \mathbb{R}_{[0,1]}$ by the Hidden Markov Model process: 
		$HMM: (\mathbb{Z^{+}}^k)^t \rightarrow h_g$, where each training example consists of $k$ instantaneous data points and there are $t$ training examples for gesture type $g$.

@@

		\subsubsection{Clustering}
		For each gesture type $g$ across all training examples, k-means defines a clustering function $f_g: \mathbb{R}^n \rightarrow \mathbb{Z^{+}}$.

@@

For example, using Cluster-Matching-HMM, both accelerometer and gyroscope data together
achieves 94.6\% on all gesture types, followed by gyroscope data only at 92.3\%, and
accelerometer data only at 92.0\%.  The same general pattern is evident in the KHMM
data as well.
